---
title: "Hw2_Discrete_Optimization"
author: "yshao06"
date: "October 23, 2017"
output: pdf_document
---

# 1. Setup and Deliverables
```{r}
rm(list = ls())

# Install Package for further processing
installIfAbsentAndLoad <- function(neededVector) {
    for(thispackage in neededVector) {
        if( ! require(thispackage, character.only = T) )
        {install.packages(thispackage)}
        require(thispackage, character.only = T)
    }
}
# load required packages
needed  <-  c("optrees", "igraph", "qgraph")
installIfAbsentAndLoad(needed)
```

# 2. Matrix Representation
```{r}
# Create an adjacency matrix with 1000 x 1000 dimension
n2 <- 1000
d2 <- runif(n2*n2)
d2[d2<0.8] <- NA
d2 <- matrix(d2, nrow = n2, ncol = n2)
diag(d2) <- NA
d2[upper.tri(d2)] <- t(d2)[upper.tri(d2)]

# Create an function for converting adjacency matrix to list
AdjMatrix2List <- function(d) {
    ds <- matrix(nrow = sum(!is.na(d)), ncol = 3, dimnames = list(NULL, c("head", "tail", "weight")))
    k = 1
    for (i in 1:nrow(d)) {
        for (j in 1:ncol(d)) {
            if (!is.na(d[i, j])) {
                ds[k, 1] <- i
                ds[k, 2] <- j
                ds[k, 3] <- d[i,j]
                k = k+1
            }
        }
    }
return(ds)
}

# Use function to convert d to ds
ds2 <- AdjMatrix2List(d2)
```

# 3. Euclidean Minumum Spanning Tree
```{r}
# Create an adjacency matrix with 50x50 dimension
n3 <- 50
x3 <- round(runif(n3) * 1000)
y3 <- round(runif(n3) * 1000)
plot(x3, y3, pch = 16)
```

## 3.1 Create an adjacency matrix d by calculating the Euclidean distance between every pair of points
```{r}
d3 <- matrix(0, nrow = n3, ncol = n3)
for (i in 1:nrow(d3)) {
    for (j in 1:ncol(d3)) {
        d3[i, j] <- sqrt((y3[j]-y3[i])^2 + (x3[j]-x3[i])^2)
    }
}
```

## 3.2 Using function AdjMatrix2List, create an adjacency list from d called ds
```{r}
ds3 <- AdjMatrix2List(d3)
```

## 3.3 Calculate the minimum spanning tree using Kruskal or Prim
```{r}
ds.mst3 <- msTreePrim(1:n3, ds3)
```

## 3.4 Produce a plot visualizating minimum spanning tree
```{r}
plot.mst <- function(arcList) {
    segments(x3[arcList[, 1]], y3[arcList[, 1]], x3[arcList[, 2]], y3[arcList[, 2]],
             col = "red", lty = par("lty"), lwd = par("lwd"))
}

plot(x3, y3, pch=16,  xlab='x', ylab='y', main="Visualization of MST")
plot.mst(ds.mst3$tree.arcs)
```

# 4. Hostile Agents

## 4.1 Specify this problem as a graph optimization problem
There are n nodes in the graph, representing the agents of the intelligence service. There are also  edges connecting agents i and j who know each other. Edge weight is p(ij), denoting the probability that message passed between these agents will fall into hostile hands.
To maximize the total probability that no message is intercepted, is to find the path connecting all agents where the product of pij's is minimum.

## 4.2 How do you structure your inputs to this problem
Create an adjacency matrix of n x n dimension (symmetric for undirected graph), where the value of matrix[i, j] is NA if the pair of agents are not connected, or is log(p(ij)) if they are connected.
The rationale of taking log(p(ij)) is to converting the "minimizing product" problem into "minimizing sum" problem for MST algorithm to take over.

## 4.3 Which algorithm do you use to solve it
Prim and Kruskal are the two algorithms available under MST, whose efficiency depends on the density of the graph they handled.
Presumably, the connection (edges) among agents (nodes) is sparse for the security purpose, using Kruskal algorithm is a better choice for the benefit of efificiency.

## 4.4 What is the computational efficiency of your chosen solution
The computational efficiency of the solution is nlog(n)

# 5.  Project Scheduling

## 5.1 Represent scheduling network by associating a completion time with each node
```{r}
s.labels <- c(letters[1:10])
s.nodes <- c(90, 15, 5, 20, 21, 25, 14, 28, 30, 45)
m.nodes <- matrix(data = s.nodes, dimnames = list(s.labels, NULL))
```

## 5.2 Create matrix representing nodes and edges
```{r}
s <- matrix(c("a", "b",
              "b", "c",
              "g", "d",
              "d", "e",
              "a", "f",
              "c", "g",
              "f", "g",
              "d", "h",
              "a", "i",
              "d", "j",
              "i", "j"), 
            ncol = 2, byrow = TRUE)
# Create adjacency matrix and name row and col by node labels
s.matrix <- matrix(nrow = 10, ncol = 10, dimnames = list(s.labels, s.labels))
for (i in 1:nrow(s)) {
    s.matrix[s[i, 1], s[i, 2]] <- -as.numeric(m.nodes[s[i, 1], 1])
}
s.matrix
```

## 5.4 Earliest Start Times using
```{r}
s.ds <- AdjMatrix2List(s.matrix)
s.sp <- spTreeBellmanFord(1:length(s.labels), s.ds, directed = TRUE)
s.ES <- cbind(letters[s.sp$tree.nodes], -s.sp$distances)
```

## 5.6 Earliest Finish Times = Earliest Start Time + Node Duration
```{r}
s.EF <- cbind(s.ES[, 1], as.numeric(s.ES[, 2])+s.nodes[s.sp$tree.nodes])
```

## 5.5 Earliset Overall Project Competion Time
```{r}
start.date <- "2017-11-01"
PCT <- as.Date(start.date) + max(as.numeric(s.EF[, 2]))
PCT
```

## 5.7 Latest Finish Times computed on the transpose of the graph
```{r}
s.matrix.t <- matrix(nrow = 10, ncol = 10, dimnames = list(s.labels, s.labels))
for (i in 1:nrow(s)) {
    s.matrix.t[s[i, 2], s[i, 1]] <- -as.numeric(m.nodes[s[i, 2], 1])
    # rebuild edge weights by taking the processing time of succesors
}
s.ds.t <- AdjMatrix2List(s.matrix.t)
s.sp.t <- spTreeBellmanFord(1:length(s.labels), s.ds.t, source.node = c(5, 8, 10), directed = TRUE)
s.LF <- cbind(letters[s.sp.t$tree.nodes], s.sp.t$distances+194)
```

## 5.8 Latest Start Times = Latest Finish Time - Node Duration
```{r}
s.LS <- cbind(s.LF[, 1], as.numeric(s.LF[, 2])-s.nodes[s.sp$tree.nodes])
```

## 5.9 Slack = LF - EF = LS - ES
```{r}
s.slack <- cbind(s.ES[, 1], as.numeric(s.LS[, 2]) - as.numeric(s.ES[, 2]))
```

## Consolidated output of project scheduling:
```{r}
s.table <- matrix(data = c(s.labels, s.ES[, 2], s.EF[, 2], s.LS[, 2], s.LF[, 2], s.slack[, 2]), 
                  ncol = 6, dimnames = list(NULL, c("Labels", "ES", "EF", "LS", "LF", "Slack")))
s.table
```
Task B, C, E, H, I  have scheduling flexibility, whereas task A, F, G, J are on the critical path

## 5.10 Gantt Chart (Extra)
```{r}
barplot(rbind(as.numeric(s.LF[, 2]) - s.nodes, s.nodes), horiz = TRUE, col = c("white", "gray"),
        border = 0, las = 1, names.arg = s.labels)
# following function could not be converted to pdf file properly
# library(timevis)
# gnt <- data.frame(id = 1:10, content = s.labels, start = as.numeric(s.LF[, 2]) - s.nodes,
# end = as.numeric(s.LF[, 2]))
# timevis(gnt)
```
